SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-07 back.tex
misc changes

\noindent \textbf{Signature Table:} As shown on the right side of
figure \ref{fig:spp_update}, the Signature Table keeps track of 256
most recently accessed pages. It is meant to capture memory access
patterns within a page boundary. SPP indexes into an entry of the
Signature Table using the page number. For each entry corresponding to
a page, the table stores a `last block offset' and an `old
signature'. Last block offset is the block offset of the last memory
access of that given page. The block offset is calculated with respect
to the page boundary. The signature is a 12-bit compressed
representation of the past few memory accesses for that page. The
signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
(\,Delta\,)$$ Delta is the numerical difference between the block
offset of the current and the previous memory access. In case a
matching page entry is found, the stored signature retrieved and used
to index into the Pattern Table.  This process is illustrated in
figure \ref{fig:spp_update}.

\noindent \textbf{Pattern Table:} The Pattern Table, shown on the left
side of figure~\ref{fig:spp_update} is indexed by the signature
generated from the Signature Table.  Pattern Table holds predicted
delta patterns and their confidence estimates. Each entry indexed by
the signature holds up to 4 unique delta predictions.

@@

\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
down the program speculation path using its own prefetch suggestion.
Using current prefetch as a starting point, it re-accesses the Pattern
Table to generate further prefetches.  As illustrated in
figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
Pattern Table and updating the signature based on highest confidence
prefetch from the last iteration.  The iteration counter on which SPP
manages to predict prefetch entries in the lookahead manner is
characterized as its `depth'. While doing so, SPP also keeps
compounding the confidence in each depth. Thus as depth increases,
overall confidence keeps decreasing.

\noindent \textbf{Confidence Tracking}: As shown in
figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
hits to each signature through a counter C\textsubscript{sig}. The
number of hits for a given delta per signature are tracked using a
counter C\textsubscript{delta}.  The confidence for a given delta is
approximated through C\textsubscript{d} = C\textsubscript{delta} /
C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
calculated as the ratio of the number of prefetches which led to a
demand hit to the number of prefetches recommended in total. The range
of $\alpha$ is [0,1].  $D$ is the lookahead depth. For $d = 1$, when
SPP is in non-speculative mode, P\textsubscript{0} can be thought of
as 1.  The final P\textsubscript{d} is thresholded against prefetch
threshold (T\textsubscript{p}) to reject the low confidence
suggestions and then against a numerically bigger fill threshold
(T\textsubscript{f}) to decide whether to send the prefetch to
L2 Cache (high confidence prefetch) or Last Level Cache (low
confidence prefetch). The two thresholds were empirically set to 25
and 90 respectively, on the scale of 0 to 100.

>>>>

\noindent \textbf{Signature Table:} As shown on the left side of
Figure~\ref{fig:spp_update}, the Signature Table keeps track of 256
most recently accessed pages. It is meant to capture memory access
patterns within a page boundary. SPP indexes into an entry of the
Signature Table using the page number. For each entry corresponding to
a page, the table stores a `last block offset' and an `old
signature'. Last block offset is the block offset of the last memory
access of that given page. The block offset is calculated with respect
to the page boundary. The signature is a 12-bit compressed
representation of the past few memory accesses for that page. The
signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
(\,Delta\,)$$ Delta is the numerical difference between the block
offset of the current and the previous memory access. In case a
matching page entry is found, the stored signature retrieved and used
to index into the Pattern Table.  This process is illustrated in
Figure~\ref{fig:spp_update}.

\noindent \textbf{Pattern Table:} The Pattern Table, shown on the right
side in Figure~\ref{fig:spp_update} is indexed by the signature
generated from the Signature Table.  Pattern Table holds predicted
delta patterns and their confidence estimates. Each entry indexed by
the signature holds up to 4 unique delta predictions.

@@

\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
down the program speculation path using its own prefetch suggestion.
Using current prefetch as a starting point, it re-accesses the Pattern
Table to generate further prefetches.  As illustrated in
Figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
Pattern Table and updating the signature based on highest confidence
prefetch from the last iteration.  The iteration counter on which SPP
manages to predict prefetch entries in the lookahead manner is
characterized as its `depth'. While doing so, SPP also keeps
compounding the confidence in each depth. Thus as depth increases,
overall confidence keeps decreasing.

\noindent \textbf{Confidence Tracking}: As shown in
Figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
hits to each signature through a counter C\textsubscript{sig}. The
number of hits for a given delta per signature are tracked using a
counter C\textsubscript{delta}.  The confidence for a given delta is
approximated through C\textsubscript{d} = C\textsubscript{delta} /
C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
calculated as the ratio of the number of prefetches which led to a
demand hit to the number of prefetches recommended in total. The range
of $\alpha$ is [0,1].  $D$ is the lookahead depth. For $d = 1$, when
SPP is in non-speculative mode, P\textsubscript{0} can be thought of
as 1.  The final P\textsubscript{d} is thresholded against prefetch
threshold (T\textsubscript{p}) to reject the low confidence
suggestions and then against a numerically bigger fill threshold
(T\textsubscript{f}) to decide whether to send the prefetch to
L2 Cache (high confidence prefetch) or Last Level Cache (low
confidence prefetch). The two thresholds were empirically set to 25
and 90 respectively, on the scale of 0 to 100.

### 2018-12-07 design_arch.tex
misc changes

It can be beneficial to allow a prefetcher to speculate as deeply as
possible. Often, some useful prefetches are generated long after the
confidence of the prefetcher has fallen below a point, after which the
performance degrades, due to the increase of inaccurate prefetches. However,
in order to allow deep speculation in the prefetcher, inaccurate prefetches
must be filtered out. We propose to leverage perceptron-based learning as a
mechanism to differentiate between potentially useful deeply speculated
prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF) is
placed between the prefetcher and the prefetch insertion queue, as illustrated
in figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
polluting the higher levels of the memory hierarchy.

@@

\noindent \textbf{Recording}\\
As shown in step 2 of figure~\ref{fig:PPF_Datapath} the prefetches that make
it through the inference stage are recorded in the ``Prefetch Table''. The
prefetch table is a 1,024-entry, direct mapped structure that contains all
metadata required to re-index the perceptron entries for training.

@@

On a cache block eviction, we look up the corresponding address in the
prefetch table. If there is a valid entry with this address, the filter made a
misprediction. The block was allocated in the L2 with a prefetch request that
the filter should have categorized as a useless prefetch. Thus, the
corresponding features of the prefetch request that cause these block to be
placed in the cache are used to index the tables of weights, and those weights
are adjusted accordingly.

>>>>

It can be beneficial to allow a prefetcher to speculate as deeply as
possible. Often, some useful prefetches are generated long after the
confidence of the prefetcher has fallen below a point, after which the
performance degrades, due to the increase of inaccurate prefetches. However,
in order to allow deep speculation in the prefetcher, inaccurate prefetches
must be filtered out. We propose to leverage perceptron-based learning as a
mechanism to differentiate between potentially useful deeply speculated
prefetches and likely not-useful ones. The Perceptron Prefetch Filter (PPF) is
placed between the prefetcher and the prefetch insertion queue, as illustrated
in Figure~\ref{fig:PPF_Hierarchy}, to prevent not-useful prefetches from
polluting the higher levels of the memory hierarchy.

@@

\noindent \textbf{Recording}\\
As shown in step 2 of Figure~\ref{fig:PPF_Datapath} the prefetches that make
it through the inference stage are recorded in the ``Prefetch Table''. The
prefetch table is a 1,024-entry, direct mapped structure that contains all
metadata required to re-index the perceptron entries for training.

@@

On a cache block eviction, we look up the corresponding address in the
prefetch table. If there is a valid entry with this address, the filter made a
misprediction. The block was allocated in the L2 with a prefetch request that
the filter should have categorized as a useless prefetch. Thus, the
corresponding features of the prefetch request are used to re-index the tables 
of weights, and those weights are adjusted accordingly.

### 2018-12-07 intro.tex
Typo
misc changes

\begin{figure}[t]
\includegraphics[width=\columnwidth]{Motivation}
\caption{Impact of aggressive prefetching on performance. X-asis
  represents increasingly aggressive prefetch lookahead depth. All the
  metrics have been normalized to the least aggressive
  configuration. As can be seen, total number of prefetches increases
  at a higher rate than the number of good prefetches, leading to IPC
  reduction.}
\label{Fig:Motivation}
\end{figure}

@@

Figure~\ref{Fig:Motivation} illustrates the above scenario.  Here we
consider a state-of-the-art. lookahead prefetcher -- SPP~\cite{SPP}.
Lookahead prefetchers such as SPP provide a mechanism to speculate an
arbitrary number of references ahead of the initial triggering miss.
In SPP, a throttling confidence threshold is then used to ensure that
the lookahead stops when confidence falls too low to ensure that
prefetches are accurate.  In the figure, we remove this threshold and
instead allow the prefetcher to lookahead an of arbitrary depth from 7
to 15.  The figure depicts the behavior of the {\tt 603.bwaves\_s}
SPEC CPU 2017 benchmark.  The IPC, total number of prefetches
suggested by the prefetcher (TOTAL\_PF) and the actual useful
predictions (GOOD\_PF), all have been normalized to the least
aggressive version of the prefetcher (depth 7).  As the look-ahead
depth increases, so do useful prefetches, and hence coverage.  This
coverage, however, comes at a cost of total prefetches increasing at
an even higher rate.  This leads to cache pollution and bandwidth
contention, as is clearly visible with the reduction in IPC.  Similar
behavior was observed in 12 other SPEC CPU 2017 benchmarks.

>>>>

\begin{figure}[t]
\includegraphics[width=\columnwidth]{Motivation}
\caption{Impact of aggressive prefetching on performance. X-axis
  represents increasingly aggressive prefetch lookahead depth. All the
  metrics have been normalized to the least aggressive
  configuration. As can be seen, total number of prefetches increase
  at a higher rate than the number of good prefetches, leading to IPC
  reduction.}
\label{Fig:Motivation}
\end{figure}

@@

Figure~\ref{Fig:Motivation} illustrates the above scenario.  Here we
consider a state-of-the-art. lookahead prefetcher -- SPP~\cite{SPP}.
Lookahead prefetchers such as SPP provide a mechanism to speculate an
arbitrary number of references ahead of the initial triggering miss.
In SPP, a throttling confidence threshold is then used to ensure that
the lookahead stops when confidence falls too low to ensure that
prefetches are accurate.  In the figure, we iteratively re-tuned 
this threshold and allowed the prefetcher to lookahead an of 
arbitrary depth from 7 to 15. The figure depicts the behavior 
of the {\tt 603.bwaves\_s} SPEC CPU 2017 benchmark. The IPC, the 
total number of prefetches suggested by the prefetcher (TOTAL\_PF) 
and the actual useful predictions (GOOD\_PF), all have been normalized 
to the least aggressive version of the prefetcher. As the look-ahead
depth increases, so do useful prefetches, and hence coverage. This
coverage, however, comes at a cost of total prefetches increasing at
an even higher rate. This leads to cache pollution and bandwidth
contention, as is clearly visible with the reduction in IPC.  
%Similar behavior was observed in 12 other SPEC CPU 2017 benchmarks.

### 2018-12-07 method.tex
Condensing for space, Coverage and CrossVal Images
misc changes

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|}
    \hline
    	 Block & Configuration\\
    \hline
	 \multirow{2}{1.5cm}{CPU Core} 		& 1-8 Cores, 4 GHz\\
						& 256 entry ROB, 4-wide\\
    \hline
         \multirow{2}{2.7cm}{Private L1 DCache} & 32 KB, 8-way, 4 cycles\\
						& 8 MSHRs, LRU\\
    \hline
         \multirow{2}{2.4cm}{Private L2 Cache}  & 256 KB, 8-way, 8 cycles\\
						& 16 MSHRs, LRU, Non-inclusive\\		
    \hline
         \multirow{2}{1.7cm}{Shared LLC} 	& 2MB/core, 16-way, 12 cycles\\
						& 32 MSHRs, LRU, Non-inclusive\\
    \hline
         \multirow{3}{1.1cm}{DRAM} 		& 4 GB 1-Channel (single-core)\\ 
						& 8 GB 2-Channels (multi-core)\\ 
						& 64-bit channel, 1600MT/s\\
    \hline
    \end{tabular}
    \caption{Simulation Parameters}
    \label{tab:Sim_params}
\end{table}

@@

\subsection{Testing Under Additional Memory Constraints}
\label{Method-AdditionalMem}
The default single-core configuration simulates a 2MB LLC and a single channel
DRAM with 12.8GB/s bandwidth. We extend the simulations to include memory
constraints introduced in DPC-2. Specifically we look at the following two
variations: Low Bandwidth DRAM, where DRAM bandwidth is limited to 3.2 GB/s,
and small LLC, where the LLC size is reduced to 512 KB All the multi-core
simulations are only done in the default configuration.

@@

%This is the first time that SPEC CPU 2017 benchmark suite~\cite{SPEC2017}
%has been used to characterize and measure the prefetch performance.
We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. We use all the 20
workloads available in the SPEC CPU 2017 suite. Using the SimPoint~\cite{SimPoint}
methodology, we identified 95 different program segments of 1 Billion
instructions each.\\
%
\noindent \textit{Single-core performance:} For single-core simulations, we use 200
million instructions to warm-up the microarchitectural structures and the next
one billion instructions to do detailed simulations and collect run-time
statistics. We report the IPC speedup over the baseline of no prefetching.
The final numbers reported are the geometric mean of the weighted mean speedup 
achieved per application using the SimPoint methodology.\\
%
\noindent \textit{Multi-core performance:} For multi-application workloads, we generate
100 random mixes and another 100 mixes from the memory intensive subset of
SPEC CPU 2017. For 4-core workload, 200 Million instructions are used for warm-up
and additional 1 Billion instruction simulated for collecting statistics.
Each CPU keeps executing its workload till the last CPU completes one billion
instructions after warm-up. For collecting IPC and other data, only the first
billion instructions are considered as the region of interest.

@@

Figure~\ref{Fig:Pvals} shows all the features which are finally used, 
arranged in the increasing order of their Pearson's factor. 
As can be seen 5 out of the 9 features provide a moderate to high 
correlation, with the magnitude of P-value > 0.6. The single most 
important feature, \textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90. \\ \\
%
\textbf{Per Trace Correlation} \newline Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces. Here we give special attention to features
with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the
variation of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and 
\textit{PC XOR Depth}; across all the SPEC CPU 2017 traces. 
For simplicity, the traces are arranged in an
increasing order of contribution made by the. 
It can be seen that even features with a low
overall correlation provide useful correlation (magnitude > 0.5) for
a significant number of traces. \\ \\
%
\textbf{Trimming Features Using Cross-correlation} \newline Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

@@

We initially came up a mix of 23 features. By studying
cross correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9 in
magnitude and eliminated redundant features, reducing the feature count to 9.
Thus, in the final implementation of PPF, no two features have a high
correlation between them. This way we can be sure that each feature makes a
contribution that cannot be captured using other features.

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{|c|c|c|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid 		& 1  & Indicates a valid entry in the table\\
         Tag 		& 6  & Identifier for the entry in the table\\
         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
                    	&    & useful demand fetch\\
         Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
    \hline
        PC 		& 12 & \\
        Address 	& 24 & \\
        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Meta-data required for perceptron}\\
	PC$_i$ Hash	& 12 & \multirow{2}{1.1cm}{training}\\
        Delta 		& 7  & \\
        Confidence 	& 7  & \\
	Depth		& 4  & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 85 bits}}\\
    \hline
    \end{tabular}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}

@@

\begin{table}[h]
\begin{adjustwidth}{-0.1cm}{}
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
                                            &  \multirow{5}{0.5cm}{256}    & Valid (1 bit)  &             \\
                                             &      & Tag (16 bits)        &  \multirow{2}{0.9cm}{11008}           \\
                            Signature Table  &   & Last Offset (6 bits) &  \multirow{2}{0.5cm}{bits}  \\  
                                             &      & Signature (12 bits)  &             \\
                                             &      & LRU (6 bits)         &             \\
    \hline
                                    &  \multirow{3}{0.5cm}{512}    & $C_{sig}$ (4bits)      &\multirow{2}{0.9cm}{24576}               \\
                       Pattern Table         &   & $C_{delta}$ (4*4 bits) &  \multirow{2}{0.5cm}{bits}  \\
                                             &      & Delta (4*7 bits)       &               \\
    \hline
        \multirow{4}{1.5cm}{Perceptron\newline}     & 4096*4    & \multirow{4}{0.8cm}{5 bits}  & \multirow{3}{1.1cm}{113280}             \\
        \multirow{3}{1.2cm}{Weights}                & 2048*2    &           &  \multirow{3}{0.5cm}{bits}  \\
                                                    & 1024*2    &           &               \\
                                                    & 128*1     &           &              \\
    \hline
        Prefetch                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{85 bits}       & 87040 \\
        Table\footnotemark[1]   &           &               & bits\\
    \hline
        Reject                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{84 bits}    & 86016 \\
        Table\footnotemark[2] & & & bits\\
    \hline
        \multirow{4}{1.0cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
        \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
        \multirow{2}{1.2cm}{Register}               &                        & Last Offset (6 bits) &                               \\
                                                    &                        & Delta (7 bits)       &                               \\
    \hline
        Accuracy        & 1     & C$_{total}$       & 10 bits   \\
        Counters        & 1     & C$_{useful}$      & 10 bits   \\
    \hline
        \multirow{3}{1.5cm}{Global PC\newline}      &       & $PC_1$ (12 bits)      &           \\
        \multirow{2}{1.5cm}{~Trackers}              & 3     & $PC_2$ (12 bits)      & 36 bits   \\
                                                    &       & $PC_3$ (12 bits)      &           \\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 3,222,940 bits = 39.34 KB}}\\
    \hline
    \end{tabular}
    \caption{SPP-Perc Storage Overhead}
    \label{tab:PPF_overhead}
\end{adjustwidth}
\end{table}

>>>>

\begin{table}[]
    \centering
    \begin{adjustbox}{width=0.9\columnwidth}
    \begin{tabular}{|c|c|}
    \hline
    	 Block & Configuration\\
    \hline
	 \multirow{2}{1.5cm}{CPU Core} 		& 1-8 Cores, 4 GHz\\
						& 256 entry ROB, 4-wide\\
    \hline
         \multirow{2}{2.7cm}{Private L1 DCache} & 32 KB, 8-way, 4 cycles\\
						& 8 MSHRs, LRU\\
    \hline
         \multirow{2}{2.4cm}{Private L2 Cache}  & 256 KB, 8-way, 8 cycles\\
						& 16 MSHRs, LRU, Non-inclusive\\		
    \hline
         \multirow{2}{1.7cm}{Shared LLC} 	& 2MB/core, 16-way, 12 cycles\\
						& 32 MSHRs, LRU, Non-inclusive\\
    \hline
         \multirow{3}{1.1cm}{DRAM} 		& 4 GB 1-Channel (single-core)\\ 
						& 8 GB 2-Channels (multi-core)\\ 
						& 64-bit channel, 1600MT/s\\
    \hline
    \end{tabular}
    \end{adjustbox}
    \caption{Simulation Parameters}
    \label{tab:Sim_params}
\end{table}

@@

\subsection{Testing Under Additional Memory Constraints}
\label{Method-AdditionalMem}
The default single-core configuration simulates a 2MB LLC and a single channel
DRAM with 12.8GB/s bandwidth. We extend the simulations to include memory
constraints introduced in DPC-2. Specifically we look at the following two
variations: Low Bandwidth DRAM, where DRAM bandwidth is limited to 3.2 GB/s,
and small LLC, where the LLC size is reduced to 512 KB. All the multi-core
simulations are only done in the default configuration.

@@

%This is the first time that SPEC CPU 2017 benchmark suite~\cite{SPEC2017}
%has been used to characterize and measure the prefetch performance.
%We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
We use all the 20 workloads available in the SPEC CPU 2017 suite~\cite{SPEC 2017}. Using the SimPoint~\cite{SimPoint}
methodology, we identified 95 different program segments of 1 Billion
instructions each.\\
%
\noindent \textit{Single-core performance:} For single-core simulations, we use 200
million instructions to warm-up the microarchitectural structures and the next
one billion instructions to do detailed simulations and collect run-time
statistics. We report the IPC speedup over the baseline of no prefetching.
The final numbers reported are the geometric mean of the weighted mean speedup 
achieved per application using the SimPoint methodology.\\
%
\noindent \textit{Multi-core performance:} For multi-application workloads, we generate
100 random mixes and another 100 mixes from the memory intensive subset of
SPEC CPU 2017. For 4-core workload, 200 Million instructions are used for warm-up
and additional 1 Billion instruction simulated for collecting statistics.
Each CPU keeps executing its workload till the last CPU completes one billion
instructions after warm-up. For collecting IPC and other data, only the first
billion instructions are considered as the region of interest.

@@

Figure~\ref{Fig:Pvals} shows all the features which are finally used, 
arranged in the increasing order of their Pearson's factor. 
As can be seen 5 out of the 9 features provide a moderate to high 
correlation, with the magnitude of P-value > 0.6. The single most 
important feature, \textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90. \\ \\
%
\textbf{Per Trace Correlation} \newline Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces. Here we give special attention to features
with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the
variation of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and 
\textit{PC XOR Depth}; across all the SPEC CPU 2017 traces. 
For simplicity, the traces are arranged in an
increasing order of contribution made by the feature. 
It can be seen that even features with a low
overall correlation provide useful correlation (magnitude > 0.5) for
a significant number of traces. \\ \\
%
\textbf{Trimming Features Using Cross-correlation} \newline Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

@@

We initially came up a set of 23 features. By studying
cross correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9 in
magnitude and eliminated redundant features, reducing the feature count to 9.
Thus, in the final implementation of PPF, no two features have a high
correlation between them. This way we can be sure that each feature makes a
contribution that cannot be captured using other features.

@@

\begin{table}[ht]
    \centering
    \begin{adjustbox}{width=0.9\columnwidth}
    \begin{tabular}{|c|c|c|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid 		& 1  & Indicates a valid entry in the table\\
         Tag 		& 6  & Identifier for the entry in the table\\
         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
                    	&    & useful demand fetch\\
         Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
    \hline
        PC 		& 12 & \\
        Address 	& 24 & \\
        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Meta-data required for perceptron}\\
	PC$_i$ Hash	& 12 & \multirow{2}{1.1cm}{training}\\
        Delta 		& 7  & \\
        Confidence 	& 7  & \\
	Depth		& 4  & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 85 bits}}\\
    \hline
    \end{tabular}
    \end{adjustbox}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}

@@

\begin{table}[h]
\begin{adjustwidth}{}{}
    \centering
    \begin{adjustbox}{width=0.8\columnwidth}
    \begin{tabular}{|c|c|c|c|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
                                            &  \multirow{5}{0.5cm}{256}    & Valid (1 bit)  &             \\
                                             &      & Tag (16 bits)        &  \multirow{2}{0.9cm}{11008}           \\
                            Signature Table  &   & Last Offset (6 bits) &  \multirow{2}{0.5cm}{bits}  \\  
                                             &      & Signature (12 bits)  &             \\
                                             &      & LRU (6 bits)         &             \\
    \hline
                                    &  \multirow{3}{0.5cm}{512}    & $C_{sig}$ (4bits)      &\multirow{2}{0.9cm}{24576}               \\
                       Pattern Table         &   & $C_{delta}$ (4*4 bits) &  \multirow{2}{0.5cm}{bits}  \\
                                             &      & Delta (4*7 bits)       &               \\
    \hline
        \multirow{4}{1.5cm}{Perceptron\newline}     & 4096*4    & \multirow{4}{0.8cm}{5 bits}  & \multirow{3}{1.1cm}{113280}             \\
        \multirow{3}{1.2cm}{Weights}                & 2048*2    &           &  \multirow{3}{0.5cm}{bits}  \\
                                                    & 1024*2    &           &               \\
                                                    & 128*1     &           &              \\
    \hline
        Prefetch                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{85 bits}       & 87040 \\
        Table\footnotemark[1]   &           &               & bits\\
    \hline
        Reject                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{84 bits}    & 86016 \\
        Table\footnotemark[2] & & & bits\\
    \hline
        \multirow{4}{1.0cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
        \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
        \multirow{2}{1.2cm}{Register}               &                        & Last Offset (6 bits) &                               \\
                                                    &                        & Delta (7 bits)       &                               \\
    \hline
        Accuracy        & 1     & C$_{total}$       & 10 bits   \\
        Counters        & 1     & C$_{useful}$      & 10 bits   \\
    \hline
        \multirow{3}{1.5cm}{Global PC\newline}      &       & $PC_1$ (12 bits)      &           \\
        \multirow{2}{1.5cm}{~Trackers}              & 3     & $PC_2$ (12 bits)      & 36 bits   \\
                                                    &       & $PC_3$ (12 bits)      &           \\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 3,222,940 bits = 39.34 KB}}\\
    \hline
    \end{tabular}
    \end{adjustbox}
    \caption{SPP-Perc Storage Overhead}
    \label{tab:PPF_overhead}
\end{adjustwidth}
\end{table}

### 2018-12-07 related.tex
Mostly related to images, methodology, typos etc
Added description of DA-AMPM
misc changes

Spatial prefetchers include such well-understood examples as the next-line (or
next-$n$-line) prefetcher, and the stream prefetcher, and are distinguished by prefetching
data without regard for the order in which the data will be accessed.
In addition to simpler these, Somogyi \textit{et al.} 
propose Spatial Memory Streaming (SMS)~\cite{SMS}.  SMS works by learning
the spatial footprint of all data used by a program within a region of memory,
and when that region is revisited by the program, the same spatial footprint
is prefetched.  Ishii \textit{et al.} propose the Access Map Pattern Matching 
prefetcher (AMPM)~\cite{AMPM}, which creates a map of all accessed lines within a region
of memory, and then analyzes this map on every access to see if any 
fixed-stride pattern can be identified and prefetched that is centered on 
the current access. Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{Sandbox},
which analyzes candidate fixed-offset prefetchers in a sandboxed environment
to determine which is most suitable for the current program phase.  Michaud
proposes the Best-Offset Prefetcher~\cite{BOP}, which determines the optimal
offset to prefetch while considering memory latency and prefetch timeliness.

@@

Unlike spatial prefetchers, lookahead prefetchers take program order into
account when they make predictions.  Shevgoor \textit{et al.} propose the Variable 
Length Delta Prefetcher (VLDP)~\cite{VLDP} that correlates histories of
deltas between cache line accesses within memory pages with the next delta
within that page.  SPP~\cite{SPP} and KPC's prefetching component~\cite{KPC} are
more recent examples of lookahead prefetchers.  They try to predict not only what
data will be used in the future, but also the precise order in which the data
will be used, within a given page.  Predictions made by lookahead prefetchers can
be fed back into their prediction mechanisms to predict even further down 
a speculative path of memory acesses. These prefetchers can also generalize their
learned patterns from one page, and use those patterns to make predictions in 
other pages.

>>>>

Spatial prefetchers include such well-understood examples as the next-line (or
next-$n$-line) prefetcher, and the stream prefetcher, and are distinguished by prefetching
data without regard for the order in which the data will be accessed.
In addition to simpler these, Somogyi \textit{et al.} 
propose Spatial Memory Streaming (SMS)~\cite{SMS}. SMS works by learning
the spatial footprint of all data used by a program within a region of memory,
and when that region is revisited by the program, the same spatial footprint
is prefetched. Ishii \textit{et al.} propose the Access Map Pattern Matching 
prefetcher (AMPM)~\cite{AMPM}, which creates a map of all accessed lines within a region
of memory, and then analyzes this map on every access to see if any 
fixed-stride pattern can be identified and prefetched that is centered on 
the current access. DRAM-Aware AMPM (DA-AMPM)~\cite{DA_AMPM}
is a variant of AMPM that delays some prefetches so they can be issued 
together with others in the same DRAM row, increasing bandwidth utilization.
Pugsley \textit{et al.} propose the Sandbox Prefetcher~\cite{Sandbox},
which analyzes candidate fixed-offset prefetchers in a sandboxed environment
to determine which is most suitable for the current program phase. Michaud
proposes the Best-Offset Prefetcher~\cite{BOP}, which determines the optimal
offset to prefetch while considering memory latency and prefetch timeliness.

@@

Unlike spatial prefetchers, lookahead prefetchers take program order into
account when they make predictions. Shevgoor \textit{et al.} proposed the Variable 
Length Delta Prefetcher (VLDP)~\cite{VLDP} that correlates histories of
deltas between cache line accesses within memory pages with the next delta
within that page. SPP~\cite{SPP} and KPC's prefetching component~\cite{KPC} are
more recent examples of lookahead prefetchers. They try to predict not only what
data will be used in the future, but also the precise order in which the data
will be used, within a given page. Predictions made by lookahead prefetchers can
be fed back into their prediction mechanisms to predict even further down 
a speculative path of memory acesses. These prefetchers can also generalize their
learned patterns from one page, and use those patterns to make predictions in 
other pages.
